Contextual Identity: A Framework for Usable Privacy

### 2013-02-20 paper.tex
First cut at shrinking paper for w2sp. Remove definitions section and references to it.

\section{Definitions}

Suppose:
  \begin{itemize}
    \item $R$ is a set of actors and their relationships to the user. $R$ may include people that the user knows, service providers, or network operators.
    \item $T$ is a set of tasks and their side-effects. Examples of tasks
include page navigation, posting to social networks and searching. Examples of
side-effects include cookies, geolocation data and browser history.
    \item $p$ is an impression that the user wishes to project, or that actors
in $R$ perceive. $p$ encompasses such prosaic information as email identifiers
or session identifiers, or such sentiment as ``Alice is awesome!''
  \end{itemize}

\begin{definition} We define a contextual identity $I = \{R,T,p\}$, where a user performs tasks
$T$ in view of or interaction with actors in $R$, while projecting impression
$p$.
\end{definition}

A user may have multiple contextual identities $\{I\}$. These identities change
over time, as users form and dissolve relationships, and focus on different
sets of tasks. The user's goal is to accomplish tasks in $T$,
interact with those in $R$ and project $p$ while being able to predict and
assert control over as much of $I$ as possible.

Our definition focuses on relationships because most privacy concerns stem
from causing an unexpected exchange of data with people the user
knows~\cite{wang}. Tasks are key because users are task-oriented, and it is
easy to map user actions (page navigation, sending email, posting, filling
forms) to tasks.

\begin{definition}
A contextual identity violation occurs when the user cannot predict or control
parameters $R,T,p$.
\end{definition}
When a user's tasks are shared with the wrong set $R$, the wrong projected
identifier $p$ is employed, or a user is perceived as performing tasks $T$ they
do not anticipate, a violation of the integrity of contextual identity occurs.

\begin{definition}
Linkage of contextual identities occurs when a set of $\{I\}$ is associated
with the same user.
\end{definition}

When this linkage happens against the will or without the knowledge of the
user, a violation of the integrity of contextual identity occurs.

Using these definitions, we characterize different types of violation by what
part of the contextual identity the user cannot control, and then show how
control can be reasserted by a user who employs various mitigation tools.

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
(Lack of control over $R$).\\
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private
messages, before the advent of ``Like'' and comment buttons, to the attention
of an audience that the user never anticipated~\cite{timeline}.

In all these examples, the audience $R$ of the contextual identity expanded
without user intervention.

\subsection{Unsatisfiable policies}
\label{sec:policies}
(Lack of control over $p$).\\
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult. 
Linking a gaming identity to a professional identity has already caused
problems for 2012 Maine Senate candidate and World of Warcraft gamer Colleen
Lachowicz~\cite{maine}.

It is impossible for users who want to isolate multiple contextual identities
to participate in these networks without violating the terms of service.  The
``Real Names'' policy prohibits users from asserting control over $p$.

\subsection{User interface errors}
(Lack of control over $R$).\\
Even in the absence of changes made by service providers to re-contextualize
personal information, it is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{Federated login}
\label{sec:login}
(Lack of flexibility in $p$).\\
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Similar to section~\ref{sec:policies}, many federated login systems simply
don't have the ability to allow users multiple $p$ values, or identifiers.
Other federated login systems have support for multiple identifiers, but
this feature can be poorly implemented or difficult to discover, leading to
low use.
Some login platforms and protocols, such as OAuth and BrowserID,
do not suffer from this policy or design error~\cite{browserid,oauth},
giving a user better control over which value of $p$ to employ.

\subsection{Third party tracking}
\label{sec:tracking}
(Lack of control over $T$, and correlation across values of $p$).\\
Even if a user does not participate in social networks, they are still subject
to tracking by virtue of browsing the web. Third-party cookies, flash cookies,
web bugs, device fingerprinting, geolocation, history sniffing and more are
techniques that a dedicated third party could use to collect browsing history
and link contextual identities through correlation of tasks in $T$ for various
values of $p$.

@@

\subsection{Automatic mitigations for information leakage}
Given a set of relationships $R$ and tasks $T$, one can categorize a user's
contextual identities and prevent information leakage between them.  For
example, preventing third-party cookies from being read and set across task
boundaries means that targeted ads for dating would not follow the user when
visiting a technology news site. Knowing the user's current contextual identity
also allows the browser to suppress URL bar suggestions that aren't relevant to
that identity, protecting the user from both shoulder-surfing and distraction.

Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, $R$ when
using private browsing mode may be other household members, or it may be a
service to which the user wants to limit cookie leakage.
%this requires more explanation, omitting for now. A strong contextual
%identity mechanism allows users control over $R$ or over $T$ when $R$ is
%clearly identified to them, but the various privacy modes in different web
%browsers either allows the user control over $T$ or over $R$, but not both at
%the same time.

Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}. In this case, when the
user decides to register for a service using Persona, the task $T$ they are
performing is authentication, the relationships $R$ are ones they will engage
in when using the service, and $p$ is tied to the email identifier they use to
register for the service. Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.  Persona allows the user direct
control over $p$ given a context {$R,T$} in which they're authenticating.

\subsection{Expiring contextual identities}
Oftentimes a user may have a long-term task, such as wedding planning or house
buying, that requires constructing a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
in $T$ such as cookie data.

\subsection{Recovering from Mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, a frequent cause
of regret~\cite{wang}.  Redaction allows the user to assert
better control over $T$ as well as $p$. Although this approach does not help when the service
provider decides to change the audience on the user's behalf (loss of control
over $R$), auditing gives the user a good opportunity to review posts that
might cause regrets. In the case of high-volume users, such a tool could
incorporate sentiment analysis to aid the user in finding particularly
problematic content.  This effectively helps users adjust $T$ when $R$ changes
unexpectedly.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via techniques in Section~\ref{sec:tracking}. Collusion is a
tool for visualizing cookies~\cite{collusion}, in particular third-party
cookies set by tracking sites which are typically ad networks.  Once the user
knows which sites are tracking them, they can choose to block cookies from
those sites.  This type of tool allows users good control over $R$ by limiting
who sees the side-effects in $T$.

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social
networks should not either~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities or assert better control over $p$ would be to provide applications that expire posts after some time.
Deletion and expiration also reduce $R$ especially in the case of public posts
and $T$.
%This would help users better contain the whole context ($R,T$) by segmenting
%one large audience and one large set of tasks into smaller temporal partitions.

@@

\subsection{Preventing Mistakes}
\subsubsection{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience $R$.

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks
that the user does not want to be tracked. As adoption increases, ignoring this
signal will lead to loss of reputation for service providers that explicitly
act against their users' wishes, helping
users assert control over the effects of their actions $T$.

The Google Sharing project allows users of Google services to confound the way
Google identifies them~\cite{googlesharing}. In this case $R$ contains Google
and the Google Sharing proxy. The effect is that users employ a contextual
identity with a shared $p$. These users employ it only for Google sites not
requiring login and can accomplish tasks while associating all their actions
with a random $p$ also associated with many other users.  This devalues $p$ and
makes it difficult for Google to correlate tasks with a particular user.

Channel encryption systems like TOR\footnote{The Onion Router: \url{https://www.torproject.org}} allow the users more network secrecy; this allows them to better control their audience $R$ by hiding their actions from network operators.

>>>>

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private
messages, before the advent of ``Like'' and comment buttons, to the attention
of an audience that the user never anticipated~\cite{timeline}.

\subsection{Unsatisfiable policies}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult. 
Linking a gaming identity to a professional identity has already caused
problems for 2012 Maine Senate candidate and World of Warcraft gamer Colleen
Lachowicz~\cite{maine}.

It is impossible for users who want to isolate multiple contextual identities
to participate in these networks without violating the terms of service.  The
``Real Names'' policy prohibits users from asserting control over their identity.

\subsection{User interface errors}
Even in the absence of changes made by service providers to re-contextualize
personal information, it is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{Federated login}
\label{sec:login}
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Similar to section~\ref{sec:policies}, many federated login systems simply
don't have the ability to allow users multiple identifiers.
Other federated login systems have support for multiple identifiers, but
this feature can be poorly implemented or difficult to discover, leading to
low use.
Some login platforms and protocols, such as OAuth and BrowserID,
do not suffer from this policy or design error~\cite{browserid,oauth},
giving a user better control over which contextual identity to assert.

\subsection{Third party tracking}
\label{sec:tracking}
Even if a user does not participate in social networks, they are still subject
to tracking by virtue of browsing the web. Third-party cookies, flash cookies,
web bugs, device fingerprinting, geolocation, history sniffing and more are
techniques that a dedicated third party could use to collect browsing history
and link contextual identities through correlation of tasks.

@@

\subsection{Automatic mitigations for information leakage}
Given a set of relationships and tasks, one can categorize a user's
contextual identities and prevent information leakage between them.  For
example, preventing third-party cookies from being read and set across task
boundaries means that targeted ads for dating would not follow the user when
visiting a technology news site. Knowing the user's current contextual identity
also allows the browser to suppress URL bar suggestions that aren't relevant to
that identity, protecting the user from both shoulder-surfing and distraction.

Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, the threat
model for using private browsing mode may be other household members, or it may
be a service to which the user wants to limit cookie leakage.

Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\subsection{Expiring contextual identities}
Oftentimes a user may have a long-term task, such as wedding planning or house
buying, that requires constructing a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
such as cookie data.

\subsection{Recovering from Mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, a frequent cause
of regret~\cite{wang}. 
 Although this approach does not help when the service
provider decides to change the audience on the user's behalf, auditing gives the user a good opportunity to review posts that
might cause regrets. In the case of high-volume users, such a tool could
incorporate sentiment analysis to aid the user in finding particularly
problematic content.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via techniques in Section~\ref{sec:tracking}. Collusion is a
tool for visualizing cookies~\cite{collusion}, in particular third-party
cookies set by tracking sites which are typically ad networks.  Once the user
knows which sites are tracking them, they can choose to block cookies from
those sites.

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social
networks should not either~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.
Deletion and expiration also reduce the likely audience especially in the case of public posts

@@

\subsection{Preventing Mistakes}
\subsubsection{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience.

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks
that the user does not want to be tracked. As adoption increases, ignoring this
signal will lead to loss of reputation for service providers that explicitly
act against their users' wishes, helping
users assert control over the effects of their actions.
